---
title: "Generalised Linear Model"
format: html
editor: visual
---

## DAS Group 15

### Libraries

Libraries to be used during project

```{r}
library(tidyverse) 
library(moderndive) 
library(gapminder) 
library(sjPlot) 
library(stats) 
library(jtools)
```

### Data Tidying

Here we will do some data tidying to take care of NA values

```{r}
data = read.csv("dataset15.csv")
clean_data <- na.omit(data)
## Changing categorical variables to factor
clean_data$Qualityclass <- factor(clean_data$Qualityclass)
clean_data$harvested <- factor(clean_data$harvested)
aic_list <- list()
head(clean_data)
```

### GLM - Country of origin

Here we will try a GLM with the country of origin as the only explanatory variable

```{r}
model_CO <- glm(Qualityclass ~ country_of_origin, data = clean_data,family = binomial(link = "logit"))
aic_list$modelCO <- AIC(model_CO)
```

```{r}
model_CO %>%   
  summary() 
```

### GLM - Aroma

Here we will try a GLM with the Aroma as the only explanatory variable

```{r}
model_aroma <- glm(Qualityclass ~ aroma, data = clean_data,family = binomial(link = "logit"))
aic_list$modelaroma <- AIC(model_aroma)
```

```{r}
model_aroma %>%   
  summary() 
```

### GLM - Flavor

Here we will try a GLM with the Flavor as the only explanatory variable

```{r}
model_flavor <- glm(Qualityclass ~ flavor, data = clean_data,family = binomial(link = "logit"))
aic_list$model_flavor <- AIC(model_flavor)
```

```{r}
model_flavor %>%   
  summary() 
```

### GLM - Acidity

Here we will try a GLM with the country of origin as the only explanatory variable

```{r}
model_acidity <- glm(Qualityclass ~ acidity , data = clean_data,family = binomial(link = "logit"))
aic_list$model_acidity <- AIC(model_acidity)
```

```{r}
model_acidity %>%   
  summary() 
```

### GLM - Two defects

Here we will try a GLM with the Two defects category as the only explanatory variable

```{r}
model_2_defects <- glm(Qualityclass ~ category_two_defects, data = clean_data, family = binomial(link = "logit"))
aic_list$model_2_defects <- AIC(model_2_defects)
```

```{r}
model_2_defects %>%   
  summary() 
```

### GLM - Altitude

Here we will try a GLM with the Altitude as the only explanatory variable

```{r}

model_altitude <- glm(Qualityclass ~ altitude_mean_meters, data = clean_data,family = binomial(link = "logit"))
aic_list$model_altitude <- AIC(model_altitude)
```

```{r}
model_altitude %>%   
  summary() 
```

### GLM - Year Harvested 

Here we will try a GLM with the harvest year category as the only explanatory variable

```{r}

model_harvested <- glm(Qualityclass ~ harvested , data = clean_data,family = binomial(link = "logit"))
aic_list$model_harvested <- AIC(model_harvested)
```

```{r}
model_harvested %>%   
  summary() 
```

### GLM - Complete Model

Here we will try a GLM with all of the available variables

```{r}

model_full <- glm(Qualityclass ~ country_of_origin + aroma + flavor + category_two_defects + altitude_mean_meters + acidity + harvested , data = clean_data,family = binomial(link = "logit"))
aic_list$model_full <- AIC(model_full)
```

```{r}
model_full %>%   
  summary() 
```

### GLM - Stepwise Selected Model 

```{r}
stepwise_model <- step(model_full, direction = "both", trace = 0)
aic_list$stepwise_model <- AIC(stepwise_model)
```

```{r}
stepwise_model %>%
  summary()

```

### AIC and Model Selection

Here we compare all the AIC values for the different models.

```{r}
aic_df <- data.frame(Model = names(aic_list), AIC = as.numeric(aic_list))
aic_df <- aic_df[order(aic_df$AIC), ]
aic_df
```

```{r}
# Create the bar plot
barplot(aic_df$AIC, names.arg = aic_df$Model, 
        main = "AIC Comparison of Models",, ylab = "AIC Value",
        col = "skyblue", border = "black",
        las = 2, cex.names = 0.8)

# Rotate x-axis labels if needed
# par(las = 2) 

# Add a horizontal line at the minimum AIC value
abline(h = min(aic_df$AIC), col = "red", lwd = 2, lty = 2)

```

Clearly by the plot, the two models with best performance according to AIC are the stepwise model and the model that contain all the variables. In this case, we selected the stepwise model as the best as it contains the lowest AIC and also removes 2 variables, which makes our model simpler.
